---
title: "supertues"
author: '@tomstafford'
date: "4 March 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggrepel)
```

## Bang for your buck? Candidate Ad spend in the Democractic primary correlated with votes and delegates

Taking total ad spend by 4th of March 2020, as reported by [CNN](https://edition.cnn.com/2020/03/02/politics/super-tuesday-ad-spending/index.html), and the number of delegates and votes, as reported by [Wikipedia](https://en.wikipedia.org/wiki/Results_of_the_2020_Democratic_Party_presidential_primaries#Super_Tuesday) by the end of 4th of March - "Super Tuesday" - we can look and see how directly ad spending correlates with voter preferences.


```{r data, echo=FALSE,warning=FALSE}

df =read.csv('supertuesday.csv')
df$delegates <- as.numeric(as.character(df$delegates1546))
df$votes  <- as.numeric(as.character(df$votes1546))

df <- df %>% select (-c(delegates0820,votes0820,delegates1546,votes1546))

print(df)
```

## Spend v Delegates

```{r v_delegates, echo=FALSE,warning=FALSE,message=FALSE,results=FALSE}

p <- ggplot(data = df, mapping = aes(x = spend, y=delegates,label = candidates))
p + geom_point(color='red',size=3) +
  labs(x = "Total ad spend ($million)", y = "delegates won by SuperTuesday") + 
  geom_text_repel()
  xlim(0,750)

ggsave('delegates.png')    
  

```

Safe to conclude there is no strong relationship

## Spend v Votes


```{r v_votes, echo=FALSE,warning=FALSE,message=FALSE}


p <- ggplot(data = df, mapping = aes(x = spend, y=votes, label=candidates))
p + geom_point(color='blue',size=3) +
  labs(x = "Total ad spend ($million)", y = "votes won by SuperTuesday") +
  geom_text_repel() +
  scale_y_continuous(labels = scales::comma)

  
ggsave('votes.png')  

```

Again, it is safe to conclude there is no strong relationship

## Price for votes

Let's work out how much each candidate payed per vote/delegate (note, this makes the very dubious assumption that with 0 ad spend they would get 0 votes/delegates)

```{r price_per, echo=FALSE,warning=FALSE,message=FALSE}

df <- df %>% mutate(per_delegate=(spend*1000000)/delegates,per_vote=(spend*1000000)/votes) %>% drop_na()

print(df)

p <- ggplot(data=df,
            mapping=aes(x = per_vote,y=reorder(candidates,per_vote)))

p + geom_point(size=3) +
  labs(x= "Price per vote ($)",y="") +
  scale_x_continuous(labels = scales::comma)

ggsave('pervote.png')

```

## Discussion

Enjoy discussion of this analysis around the twitter [thread here](https://twitter.com/tomstafford/status/1235136079226966016)

Caveat #1: Initial graphing done while results were still coming in, so the actual figures changed (but the general pattern held).

Caveat #2: [Apparently Bloomberg's ads were terrible](https://twitter.com/broonstunes/status/1235209867079626753) so better ads might have had an effect.

Caveat #3: Also, primary voters are almost by defintion high engagement, [so this doesn't test the case of influencing decisions from low-engagement voters](https://twitter.com/broonstunes/status/1235209867079626753)

Caveat #4: [Maybe we should think of Bloomberg's spend as a hedge against the collapse of the Biden campaign](https://twitter.com/PavelDAtanasov/status/1235224962920828929)

Caveat #5: Maybe Bloomberg was buying something other than votes (e.g. press)

Caveat #6: [Maybe it was effective, since he overtook Biden in the polls at one point (from a lower base)](https://twitter.com/mathewe/status/1235168728062529536)

Caveat #7: Bloomberg's campaign stalled after disappointing debate performance, ["there's only so much advertising can do with a bad product"](https://twitter.com/mathewe/status/1235168728062529536)

## Colophon

Normally I'd do this with Python, but I was inspired by Kieran Healy's book to use R. It's a great book:

* Healy, K. (2018). [Data visualization: a practical introduction](https://socviz.co/). Princeton University Press.

* This page was useful in publishing to github [How to publish project online?](https://rpubs.com/cathydatascience/518692) by Cathy Gao, August 7, 2019. Thanks Cathy!

Repo for this analysis (the markdown file which generated this page, data and plot files) is here: [github.com/tomstafford/supertues](https://github.com/tomstafford/supertues)



